---
title: "Nitrogen_analysis_Final"
author: Patty Park
format: html
editor: visual
---

## Background

Nutrients are the essential bases for plant to be able to grow. Nitrogen, phosphorus, and potassium are a few of the necessary nutrients that plants need in order to survive. The issue is when too much of these nutrients are inputted into the environment. This threatens the ecosystem such as trophic levels, increase of invasive species, and increase disease [^1]. One particular ecosystem, vernal pools are more sensitive to these changes. Vernal pools are seasonal wetlands, meaning they are full of water during the winter season and are empty and dry during the summer season. About 90% of wetlands have been lost from urban planning and the expansion of infrastructure in these location.

One of the issues with urban expansion is the changes in natural drainage in locations that include vernal pools in these areas. Runoff may be increasing or decreasing the amount of water that end up in these pools. One other concern is that vernal pools that are close to urban location may have an increase concentration of nutrients. One of the reasoning is that it is collecting the nutrients from nearby yards with a constant application of an influx of nutrients.

Before, I had done a small project concerning the nitrogen concentration found in plant and soil matter. My hypothesis looked at vernal pool's proximity to urban spaces and the amount of nitrate as well as ammonium concentrations in these pools and how that affected the amount of nonnative plant percent cover in these areas. I had theorized that the vernal pools that were close or even had a direct pipe of runoff into these pools had an increase amount of nitrate and ammonium concentrations in it. These were the relationships as well as the hypothesis that I came up with: - Overall, the relationship between nonnative percent cover and nutrients were as followed:

$$\text{Nonnative Percent Cover} = \beta_0 + \beta_1 \text{nutrients} + \epsilon$$

-   To refine this equation to my project's goals, I rearranged the equation to include nitrate and ammonium concentration depending on if they came from plant matter or soil matter: $$\text{Nonnative Percent Cover} = \beta_0 + \beta_1 \text{nitrate and ammonium conc. soil} + \beta_2 \text{nitrogen conc. plant} + \epsilon$$

-   My hypothesis testing included the general syntax that I used when statistically analyzing my data: (ask for help: that I did this instead by looking at the normallity of my data. is it possible to do this with non native species impacted by runoff or no runoff?)

$$H_{0}: \mu_{Nonnative plant cover} - \mu_{Native plant cover} = 0$$

$$H_{A}: \mu_{Nonnative plant cover} - \mu_{Native plant cover} \neq 0$$

In this project, I was not able to find a definitive relationship between the nitrogen concentration and nonnative plant cover in my original analysis.

However, for this blog, I am going to take a step back and see if there are any interesting trends inside my data. For this analysis, I am going to look if there is a relationship between the location of where the samples were located and if that impacts the amount of nitrogen concentration at those areas.

As mentioned above, vernal pools are a sensitive ecosystem. Unfortunately, as there are only such few number left from its original amount, limited research has been done on these ecosystems. Few research papers have been published that cover the nutrient concentration in vernal pools and instead cover wetlands as a whole. One paper, "Litter breakdown among intermittently connected and unconnected geographically isolated Wetlands: how nutrient inputs alter wetland function"[^2] talks about the impact increased nutrients can have on the breakdown on litter. This brings up a certain questions: does litter decomposition happen at a more rapid pace depending on if the litter is at the edge of the pool or at the center of the pool?

Surprisingly, this paper highlights that pools with less disturbance contain higher nutrient content within litter matter. One theory that the authors theorize is that since nutrients are being delivered in a plentiful manner, the vegetation has no need to intake as much nutrients as possible.

With this in mind, I'll be looking at the following relationship:

$$\text{Nitrogen concentration} = \beta_0 + \beta_1 \text{distance from middle} + \epsilon$$

OR


$$\text{Nitrogen concentration} = \beta_0 + \beta_1 \text{Edge Nutrient Concentration} + \beta_2 \text{Transition Nutrient Concentration} + \beta_3 \text{Middle Nutrient Concentration} + \epsilon$$


##ASK IF THIS RELATIONSHIP IS GOOD OR IF IT IS LACKING

For hypothesis testing, I'll be looking at the following: ##I AM LOOKING AT THREE THINGS SO DO I HAVE TO MINUS THE THREE THINGS?

$$H_{0}: \mu_{Edge conc.} - \mu_{Transition conc.} - \mu_{Middle conc.} = 0$$
$$H_{0}: \mu_{Edge conc.} - \mu_{Transition conc.} - \mu_{Middle conc.} \neq 0$$


This blog post will cover my analysis of my data, if there is any relationship with the location of where the plant and soil sits in a vernal pool against nutrient concentration


## Data setup

Before starting our analysis, we need to call in the libraries we will be using
```{r}
library(tidyverse)
library(ggpubr)
library(gstat)
```

Let's go ahead and load in the datasets that we will be using.

```{r}
north_parcel <- read_csv("data/np_pc.csv")
nitrogen_soil <- read_csv("data/Nitrogen_soil_5.csv")
nitrogen_plant <- read_csv("data/Nitrogen_plant (5).csv")
```

My end goal is to get two different datasets: one with soil nutrient concentration and one with plant nutrient concentration. Both datasets will have geolocations included.

```{r}
#tidy up north_parcel dataset to have information that is relevent
north_parcel_clean <- north_parcel %>% 
  select(`ObjectID`, 
         `Vernal Pool Name or Number`, 
         `Vernal Pool Zone`, 
         `Transect Distance of Quadrat`, 
         x, 
         y) %>% 
  rename(Pool = `Vernal Pool Name or Number`, 
         Zone = `Vernal Pool Zone`,
         Distance = `Transect Distance of Quadrat`)


#tidy up nitrogen_soil dataset
nitrogen_soil_clean <- nitrogen_plant %>% 
  mutate(Zone = ifelse(Zone == "E", "edge", Zone),
         Zone = ifelse(Zone == "T", "transition", Zone),
         Zone = ifelse(Zone == "M", "bottom", Zone))


#join north_parcel_clean with soil dataset
soil_north_join <- right_join(nitrogen_soil_clean, north_parcel_clean, by = c('Pool', 'Zone', 'Distance'))
!duplicated(soil_north_join$ObjectID)
```













look into doing simple spatial statistics in analysis

[^1] Kneitel, J. M., & Lessin, C. L. (2010). Ecosystem-phase interactions: Aquatic eutrophication decreases terrestrial plant diversity in California vernal pools. Oecologia, 163(2), 461--469. https://doi.org/10.1007/s00442-009-1529-0

[^2] Smith, C. R., Golladay, S. W., Atkinson, C. L., & Clayton, B. A. (2022). Litter breakdown among intermittently connected and unconnected geographically isolated Wetlands: how nutrient inputs alter wetland function. Wetlands (Wilmington, N.C.), 42(6). https://doi.org/10.1007/s13157-022-01567-1


