---
title: "Nitrogen project (1)"
author: "Patty Park"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
#importing data into R
nitrogen_soil <- read_csv("nitrogen_soil.csv")
north_parcel <- read_csv("np_pc.csv")
north_parcel
nitrogen_soil2 <- rename(nitrogen_soil, 'nonnative' = 'Total Nonnative PC')
nitrogen_soil2
nitrogen_soil4 <- read_csv("Nitrogen_soil (4).csv")
nitrogen_soil4 <- rename(nitrogen_soil4, 'nonnative'='Total Nonnative PC')
nitrogen_soil4 <- rename (nitrogen_soil4, 'A_N'= 'Ammonium+Nitrate (ug/g*dm)')
nitrogen_soil4 <- rename (nitrogen_soil4, 'Ammonium'='Ammonium(ug/g*dm)')
nitrogen_soil4 <- rename (nitrogen_soil4, 'Nitrate'='Nitrate (ug/g*dm)')
nitrogen_soil4
nitrogen_plant <- read_csv("Nitrogen_plant (5).csv")
#nitrogen_plant <- rename(nitrogen_plant,'nitrogen'='Nitrogen weight (ug)')
nitrogen_plant <- rename(nitrogen_plant, 'nonnative'='Total Nonnative PC')
nitrogen_plant
```



```{r}
#creating nitrate graph
nitrate_graph <- ggplot(nitrogen_soil4, aes(x=Nitrate, y=nonnative)) +labs(title="Nitrate Concentration and Nonnative Percent Cover") +xlab("Nitrate (Âµg/g*dm)") +ylab("Nonnative Percent Cover (%)") +geom_point() +geom_smooth(method=lm , color="red", se=TRUE) + theme_classic()
nitrate_graph

#make column graph? x axis is zone, y axis is nitrogen concentration (nitrate and ammonium)

#+facet_wrap(~zone) this is for regression

#start google slides (intro, methods,etc, conclusion)

```

```{r}
#creating ammonium graph
ammonium_graph <- ggplot(nitrogen_soil4, aes(x=Ammonium, y=nonnative))  +labs(title="Ammonium Concentration and Nonnative Percent Cover") +xlab("Ammonium (Âµg/g*dm)") +ylab("Nonnative Percent Cover (%)") +geom_point() +geom_smooth(method=lm , color="red", se=TRUE) + theme_classic()
ammonium_graph
```

```{r}
A_N_graph <- ggplot(nitrogen_soil4, aes(x=A_N, y=nonnative))  +labs(title="Nitrate and Ammonium Concentration in Soil and Nonnative Percent Cover") +xlab("Nitrate + Ammonium (Âµg/g*dm)") +ylab("Nonnative Percent Cover (%)") +geom_point() +geom_smooth(method=lm , color="red", se=TRUE) + theme_classic() + stat_regline_equation(label.x=15, label.y=30) + stat_regline_equation(aes(label = ..rr.label..),label.x=15, label.y=28)
 # stat_cor(aes(label=..rr.label..), 
A_N_graph
```

```{r}
plant_graph <- ggplot(nitrogen_plant, aes(x=nitrogen, y=nonnative))  +labs(title="Nitrogen concentration in Plant Litter and Nonnative Percent Cover") +xlab("Nitrogen (Âµg N/ug)") +ylab("Nonnative Percent Cover (%)") +geom_point() +geom_smooth(method=lm , color="red", se=TRUE) + theme_classic()+stat_regline_equation(label.x=0.005, label.y=30) + stat_cor(aes(label=..rr.label..), label.x=0.005, label.y=28)
plant_graph
```


```{r}
#Nitrate
#normality of data (not manipulated)
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
#find residuals
fit_n <- lm(nonnative~Nitrate, data=nitrogen_soil4)
res_n<- fit_n$residuals

#run tests to test for normality
hist(nitrogen_soil4$Nitrate)
#qqPlot(nitrogen_soil4$Nitrate, main = "qqPlot of Nitrate concentration")
shapiro.test(nitrogen_soil4$Nitrate)

summary(fit_n)

#anova
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
nitrate_aov <- aov(nonnative ~ Nitrate, data=nitrogen_soil4)
summary(nitrate_aov)
```

```{r}
#manipulate data#
nitrogen_soil4$log_nitrate <- log(nitrogen_soil4$Nitrate)
nitrogen_soil4$sq_nitrate <-sqrt(nitrogen_soil4$Nitrate)
nitrogen_soil4

#histogram with manipulated data
hist(nitrogen_soil4$log_nitrate)
hist(nitrogen_soil4$sq_nitrate)
#log data seems most normal

#run test for normality with manipulated data
#qqPlot(nitrogen_soil4$log_nitrate, main = "qqPlot of log Nitrate concentration")
shapiro.test(nitrogen_soil4$log_nitrate)

#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
#find residuals
fit_n_log <- lm(nonnative~log_nitrate, data=nitrogen_soil4)
res_n_log<- fit_n_log$residuals

summary(fit_n_log)

#anova
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
nitrate_aov_log <- aov(nonnative ~ log_nitrate, data=nitrogen_soil4)
summary(nitrate_aov_log)
```
Running the two tests, data for Nitrate is not normally distributed
Looking at summary of res for Nitrate, there is no significance between the two slopes: 0.61680 (H0)
Anova test: 0.617, slope is not significant
even wih transformed data, there is no significance between two slopes

```{r}
#Ammonium
#normality of data (not manipulated)
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
#find residuals
#fit_a <- lm(nonnative~Ammonium, data=nitrogen_soil4)
#res_a<- fit$residuals

#run tests to test for normality
hist(nitrogen_soil4$Ammonium)
#qqPlot(nitrogen_soil4$Ammonium, main = "qqPlot of Ammonium concentration")
shapiro.test(nitrogen_soil4$Ammonium)


#summary(fit_a)

#anova
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
ammonium_aov <- aov(nonnative ~ Ammonium, data=nitrogen_soil4)
summary(ammonium_aov)
```

```{r}
#manipulate data#
nitrogen_soil4$log_ammonium <- log(nitrogen_soil4$Ammonium)
nitrogen_soil4$sq_ammonium <-sqrt(nitrogen_soil4$Ammonium)
nitrogen_soil4

#histogram with manipulated data
hist(nitrogen_soil4$log_ammonium)
hist(nitrogen_soil4$sq_ammonium)
#log data seems most normal

#run test for normality with manipulated data
#qqPlot(nitrogen_soil4$log_ammonium, main = "qqPlot of log Ammonium concentration")
shapiro.test(nitrogen_soil4$log_ammonium)

#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
#find residuals
fit_a_log <- lm(nonnative~log_ammonium, data=nitrogen_soil4)
res_a_log<- fit_a_log$residuals

summary(fit_a_log)

#anova
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
ammonium_aov_log <- aov(nonnative ~ log_ammonium, data=nitrogen_soil4)
summary(ammonium_aov_log)
```
No significance between the slopes. 0.9060 (H0)

```{r}
#Nitrate plus Ammonium
#normality of data (not manipulated)
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
#find residuals
#fit_a_n <- lm(nonnative~A_N, data=nitrogen_soil4)
#res_a_n<- fit$residuals

#run tests to test for normality
hist(nitrogen_soil4$A_N)
#qqPlot(nitrogen_soil4$A_N, main = "qqPlot of Nitrate plus Ammonium concentration")
shapiro.test(nitrogen_soil4$A_N)

#nitrate
#summary(fit_a_n)

#anova
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
a_n_aov <- aov(nonnative ~ A_N, data=nitrogen_soil4)
summary(a_n_aov)

```
No significance between the slopes: 0.7104 (H0)

```{r}
#manipulate data#
nitrogen_soil4$log_a_n <- log(nitrogen_soil4$A_N)
nitrogen_soil4$sq_a_n <-sqrt(nitrogen_soil4$A_N)
nitrogen_soil4

#histogram with manipulated data
hist(nitrogen_soil4$log_a_n)
hist(nitrogen_soil4$sq_a_n)
#log data seems most normal

#run test for normality with manipulated data
#qqPlot(nitrogen_soil4$log_a_n, main = "qqPlot of log Ammonium concentration")
shapiro.test(nitrogen_soil4$log_a_n)

#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
#find residuals
#fit_a_n_log <- lm(nonnative~log_a_n, data=nitrogen_soil4)
#res_a_n_log<- fit_a_n_log$residuals

#summary(fit_a_n_log)

#anova
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
a_n_aov_log <- aov(nonnative ~ log_a_n, data=nitrogen_soil4)
summary(a_n_aov_log)

```


```{r}
#nitrogen in plants
#normality of data (not manipulated)
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
#find residuals
#fit_plant_n <- lm(nonnative~nitrogen, data=nitrogen_plant)
#res_plant_n<- fit$residuals

#run tests to test for normality
hist(nitrogen_plant$nitrogen)
#qqPlot(nitrogen_plant$nitrogen, main = "qqPlot of Nitrogen in Plant Matter")
shapiro.test(nitrogen_plant$nitrogen)

#nitrate
#summary(fit_plant_n)

#anova
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
plant_n_aov <- aov(nonnative ~ nitrogen, data=nitrogen_plant)
summary(plant_n_aov)

```
```{r}
#manipulate data#
nitrogen_plant$log_nitrogen <- log(nitrogen_plant$nitrogen)
nitrogen_plant$sq_nitrogen <-sqrt(nitrogen_plant$nitrogen)
nitrogen_plant

#histogram with manipulated data
hist(nitrogen_plant$log_nitrogen)
hist(nitrogen_plant$sq_nitrogen)
#sq data seems most normal

#run test for normality with manipulated data
#qqPlot(nitrogen_plant$sq_nitrogen, main = "qqPlot of sq nitrogen concentration")
shapiro.test(nitrogen_plant$sq_nitrogen)

#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
#find residuals
#fit_plant_sq <- lm(nonnative~sq_nitrogen, data=nitrogen_plant)
#res_plant_sq<- fit_a_log$residuals

#summary(fit_plant_sq)

#anova
#H0:ðœ‡d (normality of data) = 0
#HA:ðœ‡d (normality of data) â‰  0
plant_aov_log <- aov(nonnative ~ sq_nitrogen, data=nitrogen_plant)
summary(plant_aov_log)
```
data is normally dstributed from qqplot. 


```{r}
#manipulating data for Nitrate (log and squared)
#nitrogen_soil4$log_nitrate <- log(nitrogen_soil4$Nitrate)
#nitrogen_soil4$sq_nitrate <-sqrtnitrogen_soil4$Nitrate)
#
```

```{r}
#not needed anymore
#Nitrate graph with separate zones
#nitrate_graph2 <- ggplot(nitrogen_soil3, aes(x=Nitrate, y=nonnative)) +geom_point() +geom_smooth(method=lm , color="red", se=TRUE) + theme_classic() + facet_wrap(~Zone)
#nitrate_graph2

```

``` {r}
#not needed anymore
#ammonium graph with separate zones
#ammonium_graph2 <- ggplot(nitrogen_soil3, aes(x=Ammonium, y=nonnative)) +geom_point() +geom_smooth(method=lm , color="red", se=TRUE) + theme_classic() + facet_wrap(~Zone)
#ammonium_graph2

```

```{r}
#not needed anymore
#plot(x=nitrogen_soil2$Nitrate, y=nitrogen_soil2$nonnative, ,xlab="Nitrate concentration", yxlab="Total Nonnative PC", main="Nitrate and Nonnative PC")

```